Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme\\ \red{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}

### 2020-04-27 LME-MPM.tex
Hasta Rigid Block

The predictor-corrector algorithm has
been described in the classic literature \cite{Hughes2000}, and its
stability and computational advantages were widely proven by Liu
\cite{Xiaojian94}. The ``classic'' \acrfull{npc} algorithm starts with a
predicted value of the nodal velocities at the $(k+1)$th time step, denoted by $\vec{\tilde{v}}_I^{k+1}$, which is calculated as follows:
\begin{equation}
  \label{eq:Predictor-velocity-I}
  \vec{\tilde{v}}_I^{k+1} = \vec{v}_I^k + (1 - \gamma)\ \Delta t\ \vec{a}_I^k.
\end{equation}

@@

\subsection{Local \textit{max-ent} approximants}
\label{sec:local-max-ent}
The popularity of \acrshort{mpm} has increase notoriously during
the recent years due to its ability to deal with large strain problems
without mesh distorsion issues inherent to mesh based methods like
\acrshort{fem}, see Zdzislaw \cite{Wieckowski2004}. However, in the simulations
made with the original \acrshort{mpm}, there are numerical noises when particles
crossing the cell boundaries. \acrfull{lme} approximation schemes were
first introduced by Arroyo \& Ortiz (2006)\cite{Arroyo2006} has been
recently tested under \acrshort{mpm} framework by Wobbes {\it et al.}
(2020)\cite{Wobbes2020} where they prof that simulations performed
with \acrshort{lme}, shows considerably more accurate stress
approximations for \acrshort{mpm}. Although, in \cite{Wobbes2020}
authors does not deep in how the regularization parameter $\beta$
affects to the accuracy and stability of the solution. The basic idea
of the shape functions based on such an estimate is to interpret the shape function $N_I(\vec{x})$ as a probability. This allow us to introduce two important limits:
the principle of maximum-entropy (\textit{max-ent}) statistical
inference stated by \cite{Jaynes1957}, and the Delaunay triangulation
which ensures the minimal width of the shape function.

This approximation scheme represents a optimal compromise, in the sense of
Pareto, between the \textit{unbiased statistical inference} based on
the nodal data which leads to the principle of \textit{maximum-entropy}
stated by Jaynes \cite{Jaynes1957}, and the definition of local shape
functions of \textit{least width} the least biased shape functions.

@@

dacay with distance away from the corresponding nodes is a desirable property. To reach to this objective \cite{Arroyo2006} propose the following linear program,

@@

The regularization o \textit{thermalization} parameter
between the two criterion $\beta$ has Pareto optimal values in the range

@@

Z(\vec{x}, \vec{\lambda})$. The traditional way to obtain such a minimiser is using Eq.~(\ref{eq:LME-J}) to calculate small increments of $\partial\vec{\lambda}$ in a Newton-Raphson approach. Where $\tens{J}$ is the Hessian matrix, defined by:

@@

Employing the chain rule, rearranging and considering $\beta$ as a constant, Arroyo and Ortiz~\cite{Arroyo2006} obtained the following expression for the gradient of the sahep function.

@@

10$^{-6}$, for example,  would ensure a reasonable range of
neighbours, see \cite{Arroyo2006} for details.

@@

find the neighbour nodes of a given integration point. An aditional remark is that in similariti  to alternative non-polynomial meshfree basis functions, the \acrshort{lme} approximation scheme requires more than $d+1$ nodes to determine the values of the shape functions as well as their derivatives at any point in the convex hull of the nodal set, where $d$ is the dimension of the problem.

This interpolation technique avoid one important shortcoming when using
\acrshort{gimp} or B-Spline \acrshort{mpm} regarding the computational domain boundaries
Steffen {\it et al.} (2008)\cite{Steffen2008b}. Which is the
additional considerations in the applications of the boundary
conditions. This is motivated because of their increased extents, it
is possible for particles to influence, and be influenced by, nodes
that lie outside of the simulation domain. Some researchers have solve
this problems with ``extra'', or the called ``ghost'' nodes by other investigators. Anyway this nodes requires especial treatments similar
to those employed in the \acrfull{sph}, for
further details see Liu \& Liu (2003)\cite{Liu2003}. The approach here
described does not requires the employ of this artifacts.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Particle-discretization}
  }
  \caption{This tolerance defines the limit values of the influence
    radius and is used thereafter to find the neighbour nodes of a
    given integration point. The picture also shows the neighbourhood
    criterion to select those node inside of \gls{domain}.} 
  \label{fig:Particle-discretization}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Due to the \short{fem}-compatibility, the \acrshort{lme} shape function is degenerated to
linear finite element shape function if $d+1$ neighbouring nodes are
chosen as the support. Furthermore, with a conveniently adopted
\textit{regularization} parameter it is possible to get a \acrshort{gimp}-like
shape function. A proof of this statements is observed in figure~\ref{fig:LME_MPM}.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \subfigure[Q4]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dy}
    \end{tabular}  
  }
  \subfigure[$\text{LME}_{17}$]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun_dy}
    \end{tabular}
  }
  \subfigure[uGIMP]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun_dy}
    \end{tabular}
  }
  \subfigure[$\text{LME}_{10}$]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun_dy}    
    \end{tabular}
  }
  \subfigure[$\text{LME}_{5}$]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun_dy}
    \end{tabular}
  }
  \caption{Comparative of linear piecewise shape functions (Q4) and
    \acrshort{ugimp} shape functions \textit{versus}  \acrshort{lme}
    approximation for a two-dimensional arrangement of nodes, and
    spatial derivatives for several values of $\gamma = \beta/h^2$.}
  \label{fig:LME_MPM}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this research and in \cite{Arroyo2006}, $\Beta$ is a scalar as the
influence area of the shape function is controlled by the Euclidean
norm, therefore the search area is geometrically a circle in 2D, or a
sphere in 3D. Building upon the idea of anisotropic shape functions,
\cite{Kochmann2019} introduced an enhanced version of the original
\acrshort{lme} scheme, which uses an anisotropic support to deal with 
tensile inestability. Nonetheless this is out of the scope of the
present document but will be incorporated in future research.

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the local \textit{max-ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm} Two different test has
been adopted for this purpose, the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995} and the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009}. Thought them, the accuracy of the \acrshort{npc}
scheme is compared to the standard \acrshort{fe}. In addition \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. To avoid some mesh-dependent issues, in all
calculations a regular background mesh was setted. All simulations were performed with in-hose software.

\subsection{Dyka bar \cite{Dyka1995}}
\label{sec:dyka-bar}

This benchmark was proposed due to its ability to shows the capability
of the proposed time integration algorithm to avoid velocity
fields instabilities. It consists in a one-dimensional bar with a
length of 0.1333 meters, sketched in the
figure~\ref{fig:Dyka_Bar}. The boundary conditions are: in the right
border displacement are constrained ($\vect{v} \rvert_{x=L} = 0$) and
in the left displacement are let free ($\tens{\sigma} \rvert_{x=0} =
0$). And a initial velocity of $\vect{v}_o = - 5\ m/s$ is given to the
last quarter of it. Finally, the elastic parameters consider for this test are:
\begin{itemize} 
\item  Density : $7833\ kg/m3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $200 \cdot 10^9\ Pa$
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures/Dyka_bar.tex}}
  \caption{Geometrical description of the Dyka \cite{Dyka1995} bar.}
  \label{fig:Dyka_Bar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this case a duration of 0.0001 seconds for the simulation is
consider. Therefore, the elastic wave generated travel thorough the bar
(from A to C and back to A) at least two times. For the spatial
discretization we adopt a set of seven nodal mesh sizes (0.1, 0.3325, 0.5,
1.0, 3.3325, 6.665, 10.0) in millimeters. For each element a number of
four particles was selected. In the initial layout, particles are
occupying the exact quadrature points of a linear quadrilateral. With
the exception of the \acrshort{ugimp} simulation where gaps or overlap between
voxels of each particle are not allowed. In those cases, each particles
occupy the center of each cell quarter. For all simulations, time step
is controlled by a Courant-Friedrichs-Levy condition of 0.1, were the adopted
celerity is computed as,
  \label{eq:Cel}
  Cel = \max\{\max_{p \in \Omega_p}\{ \vect{v}_p \} , \max_{p \in \Omega_p}\{ \sqrt{\frac{E_p}{\rho_p}} \} \}
A important consideration regarding modellization concerns to the
background mesh. Notice that free border of the bar has a maximum
horizontal displacement of 0.03 millimeters, therefore 
a computational domain with an extra gap of 0.03 millimeters is
required in order to accommodate the unconstrained displacement of the
particles in the left border of the bar. Naturally this problem arise
when the mesh size is small enough that relative displacement of the
particles are larger then the distance to the border, so grid crossing
phenomena could appear even in those cases with infinitesimal
displacements.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure*}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Error_evol}
  }
  \caption{Velocity evolution at the point in the Dyka bar left side,
    convergence plots for \acrshort{fe} and \acrshort{npc}. The plot is subdivided with
    colours, the darker part of the diagram shows coincides when the
    relative movement of the particles is large enough to produce the
    grid crossing phenomena. The lightest part of the diagram
    coincides when the relative movement of the particles in
    negligible in comparison with the mesh size. And in the middle
    region a transition behaviour take place.}
  \label{fig:Dyka-error-evol}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this case, an analytical solution is possible
thought the characteristics method described in the appendix
\ref{app:analytical_sol}. To measure the convergence of the solutions
for the different time integration and approximation schemes the
root-mean-square (RMS) error in the velocity field is computed. RMS
error is defined as
  \label{eq:RMS}
  RMS = \sqrt{\frac{1}{N} \sum^{N}_p \left( \vect{v}_p - \hat{\vect{v}}_p \right)^2},
where $\vect{v}_p$ and $\hat{\vect{v}}_p$ are respectively the analytical and
numerical solutions evaluated in the final time step in the position
of each particle. A first comparative between both time integration
scheme is plotted in figure \ref{fig:Dyka-NPC-FE}. It demonstrates the
superior performance of the \acrshort{npc} \textit{versus} the \acrshort{fe}. In the \acrshort{npc} the
spurious oscillations are quickly mitigated in the first time steps,
and does not propagate the error in time in opposite to \acrshort{fe} where the
simulation becomes unstable after $6E^{5}$ seconds. Figure
\ref{fig:Dyka-error-evol} also remarks the remarkable difference
between both schemes. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_01}
  }
  \caption{Comparative of the \acrshort{npc} \textit{versus} the \acrshort{fe}. In the
    picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Figure \eqref{fig:Dyka-LME-gamma} shows the sensibility of the \acrshort{lme}
approximation scheme to variations in the a-dimensional parameter
$\beta$ that controls the value of the regularization parameter
$\gamma$ depending of the mesh size. Notice how lower values of
gamma exhibits a behaviour with a soft decay in some parts of the
simulation due to the increase of nodes adopted to regularize the
solutions. This capability could be useful in simulations where
extremely noise oscillations could damage the solutions like memory
materials. On the other hand larger values of the parameter $\beta$
makes the solution tend to the linear \acrshort{fem} solution as the athermal
limit is reached \cite{Arroyo2006}. Intermediate values of the
regularization parameters give us a compromise between the both
scenarios here described. An additional observation concerning the
solution sensibility to regularization parameters occurs when mesh
size decreases. For larger mesh size where the relative particle displacement
is negligible in comparison with the cell size, the global behaviour
is \acrshort{fem}-like, therefore larger values of $\gamma$ has offers better
results. On the other hand, when mesh size is small enough to produce
grid-crossing and mesh-free behaviour is required to ensures the
convergence of the solution, tiny values of $\gamma$ has a better
performance. Convergence plot in figure \eqref{fig:Dyka-error-evol}
shows how the slope for the larger values of $\gamma$ decreases
monotonically with the value of the mesh size, in contrast to larger
values of it, whose suffers a punishment of the performance when significant
movement of the particles occurs as far mesh size is reduced.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_LME_gamma_comparative}
  }
  \caption{Sensitive of LME approximants performance to changes in the
    adimensional regularization parameter $\gamma = \beta/h^2$. To
    illustrate it, the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-LME-gamma}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Figure \ref{fig:Dyka-uGIMP-LME} compares the performance of the
\acrshort{ugimp}~\cite{Bardenhagen2004} shape function \textit{versus} the \acrshort{lme} approximation scheme with
a adimensional regularization parameter $\gamma$ of 4.0. Although it
does not shows remarkable differences, \acrshort{lme} approximats exhibits a more
robust behaviour than the \acrshort{ugimp} shape functions. Regarding this, notice
the absence of \acrshort{ugimp} values for a mesh size of 0.3325 and 0.5
millimeters. This is because during these simulation the \acrshort{ugimp} suffered an unstable
increase of the error which yield unacceptable results. A feasible
explanation for this phenomena could be the presence of numerical
cancellation which could produce gaps between voxels. Further research

>>>>

The predictor-corrector algorithm has
been described in the classic literature \cite{Hughes2000}, and its
stability and computational advantages were widely validated by Liu
\cite{Xiaojian94}. The ``classic'' \acrfull{npc} algorithm starts with a
predicted value of the nodal velocities at the $(k+1)$th time step, denoted by $\vec{\tilde{v}}_I^{k+1}$, which is calculated as follows:
\begin{equation}
  \label{eq:Predictor-velocity-I}
  \vec{\tilde{v}}_I^{k+1} = \vec{v}_I^k + (1 - \gamma)\ \Delta t\ \vec{a}_I^k.
\end{equation}

@@

\subsection{Local \textit{Max-Ent} approximants}
\label{sec:local-max-ent}
The popularity of the \acrshort{mpm} has increased notoriously during
the recent years due to its ability to deal with large strain problems
without mesh distorsion issues inherent to mesh based methods like
\acrshort{fem}, see Wi{\c{e}}ckowski \cite{Wieckowski2004}. However, in the simulations
made with the original \acrshort{mpm}, numerical noises appear when particles
cross the cell boundaries. \red{Solving this issue is the main goal of the employment of the \acrshort{LME} shape functions}.

\acrfull{lme} approximation schemes were
first introduced by Arroyo \& Ortiz (2006)\cite{Arroyo2006} and has been
recently tested under \acrshort{mpm} framework by Wobbes {\it et al.}
(2020)\cite{Wobbes2020}. \red{The performed simulations of \acrshort{mpm}
within \acrshort{lme} in \cite{Wobbes2020} shows considerably more accurate stress
approximations than traditional \acrshort{mpm} schemes. However, 
how the regularization parameter $\beta$
affects to the accuracy and stability of the solution is not assessed deeply in that research, which is considered interesting in the present manuscript; the tuning of this $\beta$ parameter allows to make the comparison of the accuracy against analogous traditional \acrshort{mpm} shape function.}

The basic idea
of the shape functions based on such an estimate is to interpret the shape function $N_I(\vec{x})$ as a probability. This allows us to introduce two important limits:
the principle of maximum-entropy (\textit{max-ent}) statistical
inference stated by \cite{Jaynes1957}, and the Delaunay triangulation
which ensures the minimal width of the shape function.

This approximation scheme represents an optimal compromise, in the sense of
Pareto, between the \textit{unbiased statistical inference} based on
the nodal data which leads to the principle of \textit{Maximum-Entropy}
stated by Jaynes \cite{Jaynes1957}, and the definition of local shape
functions of \textit{least width} the least biased shape functions.

@@

decay with distance away from the corresponding nodes is a desirable property. To reach to this objective \cite{Arroyo2006} propose the following linear program,

@@

The regularization or \textit{thermalization} parameter
between the two criterion, $\beta$, has Pareto optimal values in the range

@@

Z(\vec{x}, \vec{\lambda})$. The traditional way to obtain such a minimiser is using Eq.~(\ref{eq:LME-J}) to calculate small increments of $\partial\vec{\lambda}$ in a Newton-Raphson approach. $\tens{J}$ is defined as the Hessian matrix, obtained by:

@@

Employing the chain rule, rearranging and considering $\beta$ as a constant, Arroyo and Ortiz~\cite{Arroyo2006} obtained the following expression for the gradient of the shape function.

@@

10$^{-6}$ \red{in this research} \cancel{for example},  would ensure a reasonable range of
neighbours (see \cite{Arroyo2006} for details).

@@

find the neighbour nodes of a given integration point. An additional remark is that, analogous to alternative non-polynomial meshfree basis functions, the \acrshort{lme} approximation scheme requires more than $d+1$ nodes to determine the values of the shape functions as well as their derivatives at any point in the convex hull of the nodal set, where $d$ is the dimension of the problem.

This interpolation technique avoids important shortcomings when using
\acrshort{gimp} or B-Spline \acrshort{mpm} regarding the computational domain boundaries
(see Steffen {\it et al.} (2008)\cite{Steffen2008b}), which are related to the
additional considerations in the application of the boundary
conditions. Motivated by their increased extents, particles may share an influence radius that lies outside of the simulation domain. Some researchers have solved this problem with the so called ``extra'' or ``ghost'' nodes \cancel{by other investigators}. These nodes require especial treatment, similar
to those employed in the \acrfull{sph}, for
further details see Liu \& Liu (2003)\cite{Liu2003}. The approach here
described does not requires the employment of this artifices.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Particle-discretization}
  }
  \caption{This tolerance defines the limit values of the influence
    radius and is used thereafter to find the neighbour nodes of a
    given integration point. The picture also shows the neighbourhood
    criterion to select those node inside of \gls{domain}.} 
  \label{fig:Particle-discretization}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Due to the \short{fem}-compatibility, the \acrshort{lme} shape function is degenerated to
linear finite element shape function if $d+1$ neighbouring nodes are
chosen as the support. Furthermore, with a conveniently adopted
\textit{regularization} parameter it is possible to get a \acrshort{gimp}-like
shape function. A proof of this statements is observed in figure~\ref{fig:LME_MPM}\red{COMENTAR MAS; ES LA DIFERENCIA CON WOBBES}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \subfigure[Q4]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dy}
    \end{tabular}  
  }
  \subfigure[$\text{LME}_{17}$]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun_dy}
    \end{tabular}
  }
  \subfigure[uGIMP]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun_dy}
    \end{tabular}
  }
  \subfigure[$\text{LME}_{10}$]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun_dy}    
    \end{tabular}
  }
  \subfigure[$\text{LME}_{5}$]{
    \begin{tabular}{c}
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun_dy}
    \end{tabular}
  }
  \caption{Comparative of linear piecewise shape functions (Q4) and
    \acrshort{ugimp} shape functions \textit{versus}  \acrshort{lme}
    approximation for a two-dimensional arrangement of nodes, and
    spatial derivatives for several values of $\gamma = \beta/h^2$.}
  \label{fig:LME_MPM}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this research and in \cite{Arroyo2006}, $\Beta$ is a scalar as the
influence area of the shape function is controlled by the Euclidean
norm, therefore the search area is geometrically a circle in 2D, or a
sphere in 3D. Building upon the idea of anisotropic shape functions,
\cite{Kochmann2019} introduced an enhanced version of the original
\acrshort{lme} scheme, which uses an anisotropic support to deal with 
tensile instability. \red{This is another benefit of the proposed methodology, that, although is out of the scope of the
present document, will be incorporated in future research.}

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the Local \textit{Max-Ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm}. Two different test have
been adopted for this purpose: the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995} and the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009}. On the one hand, the accuracy of the \acrfull{npc}
scheme is compared to the standard \acrfull{fe} in order to assess the performance of the time integration schemes. On the other hand,  \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. To avoid some mesh-dependent issues, in all
calculations a regular background mesh was set. All simulations were performed with in-hose software.

\subsection{Dyka's bar \cite{Dyka1995}}
\label{sec:dyka-bar}

This benchmark was proposed \red{by Dyka~\cite{Dyka1995} since allows so study easily the capability}
of the proposed time integration algorithm to avoid velocity
field instabilities. It consists of a one-dimensional bar of a length of 0.1333 meters, sketched in the
figure~\ref{fig:Dyka_Bar}. The boundary conditions are: displacements are constrained ($\vect{v} \rvert_{x=L} = 0$) in the right
border, letting free in the rest of the body and ($\tens{\sigma} \rvert_{x=0} =
0$). An initial velocity of $\vect{v}_o = - 5\ m/s$ is given to the
left quarter of the bar. Finally, the elastic parameters considered for this test are:
\begin{itemize} 
\item  Density : $7833\ kg/m3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $200 \cdot 10^9\ Pa$
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures/Dyka_bar.tex}}
  \caption{Geometrical description of the Dyka \cite{Dyka1995} bar.}
  \label{fig:Dyka_Bar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this case, \cancel{a duration of } 0.0001 seconds \red{are considered }in the simulation. Therefore, the elastic wave generated travel through the bar
(from A to C and back to A) at least two times. For the spatial
discretization, a set of seven nodal mesh sizes (0.1, 0.3325, 0.5,
1.0, 3.3325, 6.665, 10.0 millimeters) are considered. For each element a number of
four particles was selected. In the initial layout, particles are
occupying the exact quadrature points of a linear quadrilateral, with
the exception of the \acrshort{ugimp} simulation, where gaps or overlap between
voxels of each particle are not allowed. In those cases, each particle
occupies the center of each cell quarter. For all simulations, time step
is controlled by a Courant-Friedrichs-Levy condition of 0.1, were the adopted
celerity is computed as:
\begin{equation}
  \label{eq:Cel}
  Cel = \max\{\max_{p \in \Omega_p}\{ \vect{v}_p \} , \max_{p \in \Omega_p}\{ \sqrt{\frac{E_p}{\rho_p}} \} \}.
\end{equation}
An important consideration regarding modellization concerns to the
background mesh. Notice that free border of the bar has a maximum
horizontal displacement of 0.03 millimeters, therefore 
a computational domain with an extra gap of 0.03 millimeters is
required in order to accommodate the unconstrained displacement of the
particles in the left border of the bar. Naturally this problem arises
when the mesh size is small enough that relative displacement of the
particles is larger then the distance to the border, so grid crossing
phenomena could appear even in those cases with infinitesimal
displacements.

In this case, an analytical solution can be obtained
through the characteristics method, described in the appendix
\ref{app:analytical_sol}. To measure the convergence of the solutions
for the different time integration and approximation schemes the
root-mean-square (RMS) error in the velocity field is computed. RMS
error is defined as
\begin{equation}
  \label{eq:RMS}
  RMS = \sqrt{\frac{1}{N} \sum^{N}_p \left( \vect{v}_p - \hat{\vect{v}}_p \right)^2},
\end{equation}
where $\vect{v}_p$ and $\hat{\vect{v}}_p$ are respectively the analytical and
numerical solutions evaluated in the final time step in the position
of each particle. \red{In Fig. \ref{fig:Dyka-error-evol} the evolution of the RMS is obtained for both time integration schemes. The right figure, with the \acrshort{NPC} results, shows lower values of the estimated error, denoting the higher performance of this methodology. About the spatial discretization, the \acrshort{LME} schemes show an error comparable to the obtained with the \acrshort{uGIMP}, being even lower close to the grid-crossing region.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure*}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Error_evol}
  }
  \caption{Velocity error evolution at the point A in the Dyka's bar ,
    convergence plots for \acrshort{fe} and \acrshort{npc}. The plot is subdivided with
    colours, the darker part of the diagram shows coincides when the
    relative movement of the particles is large enough to produce the
    grid crossing phenomena. The lightest part of the diagram
    coincides when the relative movement of the particles in
    negligible in comparison with the mesh size. And in the middle
    region a transition behaviour take place.}
  \label{fig:Dyka-error-evol}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A first comparative between the velocity results of both time integration
schemes is plotted in figure \ref{fig:Dyka-NPC-FE}. It demonstrates the
superior performance of the \acrshort{npc} \textit{versus} the \acrshort{fe}. In the \acrshort{npc} the
spurious oscillations are quickly mitigated in the first time steps,
and the error is not propagated in time, opposite to the results of the \acrshort{fe} scheme, where the
simulation becomes unstable after $6E^{5}$ seconds. \cancel{Figure
\ref{fig:Dyka-error-evol} also remarks the remarkable difference
between both schemes.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_01}
  }
  \caption{Comparative of the \acrshort{npc} \textit{versus} the \acrshort{fe}. In the
    picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Figure \eqref{fig:Dyka-LME-gamma} shows the sensibility of the \acrshort{lme}
approximation scheme to variations in the parameter
$\gamma$, which is the one that controls the value of the regularization parameter
$\beta$ together with of the mesh size. Notice how lower values of
gamma exhibits a behaviour with a soft decay in some parts of the
simulation due to the increase of nodes adopted to regularize the
solutions. This capability could be useful in simulations where
extremely noise oscillations could damage the solutions like memory
materials. On the other hand larger values of the parameter $\beta$
makes the solution tend to the linear \acrshort{fem} solution as the athermal limit is reached \cite{Arroyo2006}. Intermediate values of the
regularization parameters give us a compromise between the both
scenarios here described. An additional observation, concerning to the
solution sensibility depending on the regularization parameters, \red{is the behaviour of the solution depending on the decreasing of mesh
size}. For larger mesh sizes, where the relative particle displacement
is negligible in comparison with the cell size, the global behaviour
is \acrshort{fem}-like; therefore, larger values of $\gamma$ \red{may offer better results}. On the other hand, when mesh size is small enough to produce
grid-crossing meshfree behaviour is required to ensure the
convergence of the solution and tiny values of $\gamma$ \red{may lead to better
performances}. Convergence plot in figure \eqref{fig:Dyka-error-evol}
shows how the slope, for the larger values of $\gamma$, decreases
monotonically with the value of the mesh size, in contrast with larger
values of $\gamma$, \red{where the performance is punished with significant
movement of the particles as far as the mesh size is reduced.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_LME_gamma_comparative}
  }
  \caption{Sensitive of LME approximants performance to changes in the
    dimensionless regularization parameter $\gamma = \beta/h^2$. To
    illustrate it, the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-LME-gamma}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
